Scaffolder - Software for \change{Reproducible}{Manual} Genome Scaffolding.

### 2011-08-15 src/article.tex
Revise Results section

The Scaffolder software facilitates reproducibly joining nucleotide sequences
together into a single contiguous scaffolded super-sequence. Plain-text
scaffold files written in \remove{the} YAML \remove{syntax} specify how these
sequences should be joined\change{. and the}{The} scaffolder software is used
to generate the scaffold sequence from these instructions. In addition to
specifying which contigs are required, the scaffold file allows the contigs to
be edited into smaller sub-sequences or reverse complemented if necessary.\pb

The process of genome finishing may involve producing additional
oligonucleotide sequences to fill unknown regions in \change{the}{a} scaffold.
\remove{and} The Scaffolder \change{allows}{format provides for using} these
additional insert sequences to \remove{be used} to fill \remove{these} gaps.
These inserts can also be treated in the same manner as larger contig
sequences: trimmed and/or reverse complemented to match the corresponding gap
region size and orientation. \pb

@@

\change{The scaffold file specifies how sequence regions are joined together;
however, the sequences themselves are no  stored in the scaffold file and are
instead}{The nucleotide sequences used in the scaffold are} maintained as
a separate FASTA file. \remove{Nucleotide sequences in a FASTA-format file are
the standard output for a genome assembler, and these may be contigs, sets of
contigs scaffolded into larger sequences, or both. }The nucleotide sequences
\remove{stored in the FASTA file} are referenced in the scaffold using the
first word from in the FASTA header of the corresponding sequence. Maintaining
the nucleotide sequences in a separate file preserves the unedited sequence
and decouples the data from the specification of how it should be used to
produce the genome sequence. \pb

@@

The scaffold file is written using the YAML syntax and an example is shown in
Figure 1. This file illustrates the text attributes used to describe
a scaffold and how the sequences are correspondingly joined together in the
genome build. The basic layout of the scaffold file is a list of entries,
where each entry corresponds to a region of sequence in the generated scaffold
super-sequence. \add{Each scaffold file represents one scaffolded nucleotide
sequence and separate scaffolds should be defined in separate files.} \pb

@@

The first line of the scaffold file in Figure 1 begins with three dashes to
indicate the start of a YAML-formatted document. The first entry (highlighted
in green) begins with a dash character `-' to denote an entry in the YAML
list. This is a requirement of the YAML format: that each entry begins with
a dash line. The next line is indented by two spaces where whitespace is
required to group similar attributes together. The ``sequence'' tag indicates
that this entry corresponds to a sequence in the FASTA file and the following
line indicates the name of this sequence using the ``source'' tag. The first
word of the FASTA header is used to identify which sequence is selected from
the file. Together these three lines describe the first entry in the scaffold
as a simple sequence using the sequence named `sequence1'. On the right hand
side of Figure 1 this produces the first region in the scaffold, also shown in green. \pb

@@

The last entry in the scaffold, highlighted in blue, adds the sequence named
`sequence2' to the scaffold. This entry demonstrates how this sequence may be
manipulated prior to addition to the scaffold. The `start' and `stop' tags
trim the sequence to these coordinates inclusively.\remove{while} The
``reverse'' tag instructs scaffolder to also reverse complement the sequence.
In the putative scaffold shown in Figure 1 this completes the \remove{scaffold
} sequence. \pb

The final entry in the scaffold \remove{also} includes the ``inserts'' tag to
add additional regions of sequence. These inserts are also added as a YAML
list, with each insert beginning with a dash. The first insert, shown in
purple, uses similar attributes to that of a sequence entry; the reverse,
start and stop tags are used to trim and reverse complement the insert.
Similarly the `source' tag identifies the corresponding FASTA sequence as
`insert1'. The ``open'' and ``close'' tags are specific to inserts and
determine where the insert is added in the enclosing sequence. The region of
the sequence inside these coordinates is inclusively replaced by the specified
insert sequence. This is visualised in the putative scaffold in Figure 1 by
the black lines bisecting the blue sequence. \pb

>>>>

The Scaffolder software facilitates reproducibly joining nucleotide sequences
together into a single contiguous scaffolded super-sequence. Plain-text
scaffold files written in \remove{the} YAML \remove{syntax} specify how these
sequences should be joined\change{. and the}{The} scaffolder software is used
to generate the scaffold sequence from these instructions. In addition to
specifying which contigs are required, the scaffold file allows the contigs to
be edited into smaller sub-sequences or reverse complemented if necessary.
\add{Each scaffold file represents one scaffolded nucleotide sequence and
separate scaffolds should be defined in separate files.} \pb

The process of genome finishing may involve producing additional
oligonucleotide sequences to fill unknown regions in \change{the}{a} scaffold.
\remove{and} The Scaffolder \change{allows}{format provides functionality to
use} these additional insert sequences \remove{be used} to fill \remove{these}
gaps. These inserts can also be treated in the same manner as larger contig
sequences: trimmed and/or reverse complemented to match the corresponding gap
region size and orientation. \pb

@@

\change{The scaffold file specifies how sequence regions are joined together;
however, the sequences themselves are no  stored in the scaffold file and are
instead}{The nucleotide sequences used in the scaffold are} maintained as
a separate FASTA file. \remove{Nucleotide sequences in a FASTA-format file are
the standard output for a genome assembler, and these may be contigs, sets of
contigs scaffolded into larger sequences, or both. }The nucleotide sequences
\remove{stored in the FASTA file} are referenced in the scaffold using the
first word from \remove{in} the FASTA header of the corresponding sequence.
Maintaining the nucleotide sequences in a separate file preserves the unedited
sequence and decouples the data from the specification of how it should be
used to produce the genome sequence. \pb

@@

The scaffold file is written using the YAML syntax and an example is shown in
Figure 1. This file illustrates the text attributes used to describe
a scaffold and how the sequences are correspondingly joined together in the
genome build. The basic layout of the scaffold file is a list of entries,
where each entry corresponds to a region of sequence in the generated scaffold
super-sequence. \pb

@@

The first line of the scaffold file in Figure 1 begins with three dashes to
indicate the start of a YAML-formatted document. The first entry (highlighted
in green) begins with a dash character `-' to denote an entry in the YAML
list. This is a requirement of the YAML format: \remove{that} each entry
begins with a dash line. The next line is indented by two spaces where
whitespace is \change{required}{used} to group similar attributes together.
The ``sequence'' tag indicates that this entry corresponds to a sequence in
the FASTA file and the following line indicates the name of this sequence
using the ``source'' tag. The first word of the FASTA header is used to
identify which sequence is selected from the file. Together these three lines
describe the first entry in the scaffold as a simple sequence using the
sequence named `sequence1'. On the right hand side of Figure 1 this produces
the first region in the scaffold, also shown in green. \pb

@@

The last entry in the scaffold, highlighted in blue, adds the sequence named
`sequence2' to the scaffold. This entry demonstrates how this sequence may be
manipulated prior to addition to the scaffold. The `start' and `stop' tags
trim the sequence to these coordinates inclusively.\remove{while} The
``reverse'' tag \add{also} instructs scaffolder to \remove{also} reverse
complement the sequence. In the putative scaffold shown in Figure 1 this
completes the \remove{scaffold } sequence. \pb

The final entry in the scaffold \change{also includes}{uses} the ``inserts''
tag to add additional regions of sequence. These inserts are also added as
a YAML list, with each insert beginning with a dash. The first insert, shown
in purple, uses similar attributes to that of a sequence entry; the reverse,
start and stop tags are used to trim and reverse complement the insert.
Similarly the `source' tag identifies the corresponding FASTA sequence as
`insert1'. The ``open'' and ``close'' tags are specific to inserts and
determine where the insert is added in the enclosing sequence. The region of
the sequence inside these coordinates is inclusively replaced by the specified
insert sequence. This is visualised in the putative scaffold in Figure 1 by
the black lines bisecting the blue sequence. \pb
